A Computational Account of Default Mode Function\\ by Control Theory and Reinforcement Learning

### 2017-03-28 paper/paper.tex
comment in Elvis land

\subsection{The right and left TPJ: Prediction error signaling and world semantics}
Finally,
the right and left TPJ are known to have hemispheric differences
according to their cytoarchitectonic borders and gyrification pattern
\citep{seghier2013angular}.
Neuroscientific investigations on hemispheric functional specialization
have converged to the right versus left cerebral hemisphere as dominant for
attention versus language functions.
%
The anterior RTPJ is a key node of the DMN that  is central for
action initiation during externally structured tasks and
sensorimotor control by integrating supramodal stimulus-guided attention
\citep{corbetta2002control}.
Involvement of this DMN node was repeatedly reported in
multi-step action execution \citep{hartmann2005takes},
visuo-proprioceptive conflict (Balslev et al., 2005) and
multi-modal detection of sensory changes across
visual, auditory, or tactile stimulation in a multi-modal fMRI study
\citep{downar2000multimodal}.
In humans, direct electrical stimulation of the
RTPJ during neurosurgery was associated with altered perception
and stimulus awareness \citep{blanke2002neuropsychology}.
%
Importantly, the RTPJ has been shown to be intimately related to
supramodal prediction and error signaling.
It was argued that the RTPJ encodes actions and ensuing outcomes
without necessarily relating those to outcome value
\citep{liljeholm2013neural, hamilton2008action,
jakobs2009effects}.
Neural activity in the RTPJ has been argued to be responsible
for stimulus-driven attentional reallocation to
salient and surprising sources of information
as a ‘‘circuit breaker’’ that recalibrates control and maintenance systems
\citep{bzdok2013tpj, corbettashul2008}.
% (Bzdok et al., 2013; Corbetta et al., 2008).
Indeed, patients with right TPJ damage have particular difficulties
with multistep actions \citep{hartmann2005takes}.
In the face of large discrepancies between actual and previously predicted
environmental events the RTPJ acts a potential switch between
externally-oriented mind sets focussed on the
sensory world and internally-oriented mind sets focussed
on self-relevant mind-wandering of possible mental scenarios.
Transient RTPJ in humans for instance disrupted the
impact of predicted intentions of other individuals
\citep{young2010disruption},
a capacity believed to be subserved by the DMN.
The RTPJ might hence be an important player that shifts away
from the default active ‘‘internally directed’’ processes
to deal instead with immediate environmental objects and contexts.
% (TODO: slow endogeneous changes in state transitions by the policy and value matrices)

The left TPJ in turn exhibits a close topographical relationship to
Wernicke's area
involved in the comprehension or understanding of written and spoken language.
Neurological patients with damage caused to Wernicke's area
have a major impairment of language comprehension
when listening to others or reading a book.
Their speech
preserves natural rhythm and about normal syntax, yet the
voiced sentences are devoid of meaning (i.e., aphasia).
Abstracting from the typical semantic interpretations in linguistics
and neuropsychology,
the LTPJ probably mediates access to and integrating of world knowlege
to model action concepts
\citep{binder2011neurobiology, seghier2013angular}.
For instance, LTPJ lesions also entail problems in recognizing
others' pantomimed action towards objects
without obvious relation to processing any language content
\citep{varney1987locus}
% (Varney and Damasio 1987; Rothi et al. 1991; Buxbaum et al. 2005).
%
Also inner speech hinges on knowledge retrieval of statistical structure
about the physical and inter-personal world.
Multivariate volumetric analyses showed that the internal production of
formulated thought ("language of the mind") is closely related to the left TPJ
\citep{geva2011neural}.
A number of neuroimaging studies demonstrated LTPJ activity responses
to sementically aberrant words in sentences, which suggests a role
in binding a given semantic concept into an integrated whole.
Further,
episodic memory recall and imagination strongly draw on
complex world knowledge.
Isolated building blocks of world statistics probably get reassembled
in internally generated visual scenarios that
navigate present action, weigh hypothetical possibilities, and forcast the future.
%
The LTPJ may hence facilitate the automated prediction of events
by incorporating experience-derived models of the world
into ongoing action, planning, and problem solving.
% world -> policy matrix driven computation
% self -> value matrix driven computation

\section{Reinforcement Learning, a computational model for DMN function}
We now argue that the DMN is equipped with sufficient neurobiological substrates gradient
for implementing all the major components (memory, reward signaling, accumulators, random sampling, etc.)
of an efficient \textit{reinforcement learning (RL)} algorithm, and all these
components integrate in a plausible way.
Reinforcement Learning (RL) is learning to achieve a goal by interacting with an external
\textit{environment / world}. In a general RL framework, an \textit{agent}
interacts with an  environment in a trial-by-error fashion by taking
actions / decisions $a$, which trigger changes in the
\textit{state} of the environment
$s \rightarrow s'$, accompanied by a real-valued \textit{reward}
$r = r(s, a)$ (or regret!, since it can be negative) collected by the
agent. In this view, the environmental is partially controlled by
the action of the agent and the reward should be thought
of as an instantaneous satisfaction accompanying the execution of
an action, and may be delayed, only becoming substantial when accumulated over an extended period of time.
%% Here the subscript $t$ denotes time, and can be discrete or
%% continuous.
The environment is generally taken as stochastic,
is \textit{partially observed},
in the sense that only part of the current state is observed by
the agent. This unpredictability
% (in a first-order sense)
is realistic in a model which sets out
to explain DMN function. Indeed, it is known that, the more the external world is predictable,
the more mental activity becomes detached from the actual sensorium
\citep{antrobus1966studies, pope1978regulation}.
Conversely, the more the currently executed task is unknown and unpracticed,
the lower the DMN activity, i.e the tendency for stimulus-independent thoughts
\citep{filler1973daydreaming, teasdale1995stimulus}.
%
These ``offline'' processes may however contribute to optimized control of the organism.
A DMN framework incorporating reinforcement learning
can naturally embed human behavior
into the tension between exploitative action with immediate gains and
explorative action with longer-term reward schedules
\citep{dayan2008decision}.
The implication of the DMN in a diversity of human cognitive processes
can be parsimoneously recast as prediction error minimization
based on pervasive probabilistic simulations,
thus maximizing reward outcome at various time scales.
Such a purposeful optimization objective
may be solved by a stochastic approximation
based on a brain implementation of Markov Chain Monte Carlo sampling
\citep{tenenbaum2011grow}.

\subsection{Partially Observable Markov Decision Processes}
It is our view that at a sufficiently coarse scale, the brain is a physical system governed by the laws of
physics and can be locally described by Markov processes. Indeed, as noted in \cite{tegmark2016improved},
any system obeying the laws of classical physics can be accurately modeled as a Markov process as long as the time
step is sufficiently short (defining $s(t)$ --the state at time $t$-- as the position in phase space).
If the process has $memory$ such that the next state depends not only on the current state but also on some finite
number of past states,
rational probabilistic planning can reformulated
as a standard memoryless Markov process by simply expanding the
definition of the state $s$ to include elements of the past.

In artificial intelligence (AI), a popular abstract model for
multi-step decision processes in such an environment are so-called
\textit{Partially Observable Markov Decision Processes (POMDPs)}.
A POMDP models a decision process in which it is assumed that environment dynamics are determined by a Markov process,
but the agent cannot directly observe the underlying state. Instead, it tries to optimize a \textit{subjective} reward
signal which is tried alongside each transition in the environments state-space, and maintains a probability distribution over the set of possible
states, based on a set of observations and observation probabilities. This is a minimal amount of assumptions that can be made about an environment,
and is characteristic of so-called \textit{model-free RL}, the approach adapted by our work.
In contrast, \textit{model-based RL} starts off with some mechanistic assumptions about the dynamics of world.
These can be assumptions about the physical laws governing the agent's world or constraints on the state-space and transitions thereupon.
For example if a rat in a maze knows, for sure,
that staying still will produce no change in the environment, and in particular will not eventually lead to finding the food.
We formalize
a policy matrix to capture the space of possible actions that can be performed
on the environment given the current state. A value matrix
maps expected rewards to environmental objects and events.

@@

\subsection{Value functions and policies}
The behavior of the agent is governed by some kind of \textit{policy},
which maps states of the world to
a set of candidate actions to perform given this state. Starting a time $t=0$,
a policy $\pi$ generates a trajectory as follows
\begin{eqnarray*}
  \begin{split}
    S_1 &\sim p(.|S_0,A_0), \text{ collect rewards }R_0 = r(S_0, A_0)\\
    A_1 &\sim \pi(.|S_1)\\
    S_2 &\sim p(.|S_1,A_1), \text{ collect rewards }R_1 = r(S_1, A_1)\\
    \ldots\\
    A_{t} &\sim \pi(.|S_{t})\\
    S_{t+1} &\sim p(.|S_{t},A_{t}), \text{ collect rewards }R_{t} = r(S_{t}, A_{t})\\
    \ldots
  \end{split}
\end{eqnarray*}
Since an action taken now might have repercussions long into the feature, it turns out that the
quantity to optimize is not the instantaneous rewards $r(s, a)$, but a
cumulative reward estimate which takes into account the future. A standard for
modeling this accumulation is the so-called
time-discounted \textit{cumulative reward} function
\begin{equation}
  \label{eq:cumr}
  G^\pi = \sum_{t=0}^{\infty}\gamma^{t}R_t.
\end{equation}
This random variable\footnote{Random as it depends both on the environment's dynamic and the
  policy $\pi$ being played (which can be stochastic).}  measures the cumulative reward of
following a policy $\pi$.
\begin{mdframed}
  \textbf{Where is ``value'' buffered in the brain ?}
  The amygdala is known to be involved in significance evaluation) and
  the NAc is involved in reward evaluation.
\end{mdframed}
  The  goal of RL is then adapt this policy so as to  maximize the cumulative rewards. In
  the definition of $G^\pi$ above, the constant $\gamma$ $(0 \le \gamma < 1$) is the reward \textit{discount factor}.
Taking $\gamma = 0$   corresponds to a perfectly myopic agent who is solely concerned about their
immediate rewards and has no horizon for the future, which is not compatible with long-term goal planning
as implemented in the DMN. For there to be any learning at all, it is thus necessary that $0 < \gamma < 1$.
% \footnote{In which case we can take $T \rightarrow +\infty$, provided the reward function $r$ is bounded.}.
Thus the agent is positively biased towards considering rewards which are will
come soon. The reward importance hence decays exponentially
with distance in the future,
given that there is more uncertainty in the farsighted future.
The reward discount factor $\gamma$ can thus be seen as calibrating
the delay discounting behavior of the intelligent agent.
It is also important to appreciate that a stochastic policy estimation
is more advantageous in many RL settings.
For instance, a deterministic strategy in playing rock-paper-scissors can
be easily exploited and a uniformly random choice is optimal.

The agent's goal can then be formulated as finding policy $\pi$ which maximizes the
expected cumulative value of state-stat-pairs $(A_0,S_0)$, also known as the \textit{Q-value}, namely
\begin{equation}
  \label{eq:qval}
  Q^{\pi}(s,a) = \mathbb E [G^\pi|S_0=s,A_0=a,A_1 \sim \pi,A_2 \sim \pi,\ldots].
\end{equation}
\begin{mdframed}
In plain English, the Q-value $Q^{\pi}(s,a)$ is the expected / average / mean future reward
over all possible trajectories in which the agent starts off starts at with an environment in state
$s$, takes and action $a$, and then follows the policy $\pi$ thereafter to select future actions.
\end{mdframed}
%% Given that each $\pi$ correspondence to an exhaustive description of
%% behavioral patterns, $Q^{\pi}(S_t,A_t)$ corresponds to the epistemic value of
%% the set of actions given states in finite time.
By construction, maximizing the expectation in  \eqref{eq:qval} is intractable in general
due to the enormous number of possible trajectories the agent might need to
consider.
% \begin{mdframed}
%   \textbf{N.B.:}
%   If the Bellman-based computational ideas have a corresponding
%   neurobiological implementation, this provides a framework of how
%   optimal behavior could be computable in the human brain.
% \end{mdframed}
One popular simplification that removes this issue is to only consider the sub-family of
\textit{deterministic policies}, which map each state to the best single best action which
should be taken at thin state, and therefore define functions from states to actions.
\textit{Q-learning} (Watkins \& Dayan, 1992), a commonly used off-policy algorithm, uses the
\textit{greedy policy}
\begin{equation}
  \pi(s) = \argmax_{a \in \mathcal A}{Q}(s, a).
  \label{eq:qlearning}
\end{equation}
% where $\tilde{Q}(s, a)$ is the current estimate of the state-action value
% table.
The celebrated \textit{Bellman's equation} \cite{sutton1998reinforcement} then takes the simple form
\begin{equation}
  \begin{split}
    Q^\pi(s, a) &= \mathbb E_{s' \sim p(.|s,a)} [r(s,a) + \gamma \max_{a' \in \mathcal A}Q^\pi(s', a')]
    \\
    &= \mathbb E_{s' \sim p(.|s,a)} [r(s,a) + \gamma Q^\pi(s', \pi(s'))],
  \end{split}
  \label{eq:bellman}
\end{equation}
a \textit{fixed-point} equation which provides a recursive decomposition of optimal behavior.
A complicated dynamic programming optimization can be
broken into simpler sub-problems at different time points,
which has often been closely related to the medial prefrontal part of the DMN
\citep{koechlin1999role, braver2002role}.
Using the Bellman equation, each state can be associated with a certain value,
to guide action towards a better state, thus improving on the agent's current policy.
% It is a fixed-point equation for the deterministic policy $\pi$.
Notice that in the expectation in \eqref{eq:bellman}, the sampling is done over only things which
depend on the environment, and so can be learned off-policy by observing state-transitions
triggered by another behavioral policy, which can be stochastic.

@@

\section{Relation to other statistical models (Elvis)}
\subsection{Friston's Free-energy principle (refer to \ref{label:friston})}
\textbf{XXX: add tons of more examples and connnections}
\textbf{XXX: integrate aspects from section "ThE MARkOv DECISION PROBLEM" in Dayan and Daw}
\textbf{XXX: integrate aspects from WakeNsleep by Hinton}
\textbf{XXX: integrate aspects from Friston2014(Dropbox) on active inference}
\textbf{XXX: integrate aspects from Sutton/Barto book chapter}
In summary:
\begin{itemize}
  \item Friston's free-energy principle: brain as inference engine
  biological systems, including brains, must minimize the long-term average of surprise;
  \item A common point between the Free Energy principle and our framework
  is placing the minimization of suprise at the core of brain function.
  In RL, surprise minimization reduces to accurate prediction of
  rewarding outcomes in the future.
  \item Ortega et al. 2013 \citep{ortega2013thermodynamics} ``Thermodynamics as a theory of decision-making with information-processing costs''
\item Friston's critic of RL \citep{fristonAIorRL} (he proposes "active inference''): "This equation (equation 18) comes from the theory of dynamic programming,
pioneered in the 1950s by Richard Bellman and colleagues [2]. To
optimise control a~pð Þ x~ under this formulation, we have to: (i)
assume the hidden states are available to the agent and (ii) solve
Equation 18 for the value-function. Solving for the value-function
is a non-trivial problem and usually involves backwards induction
or some approximation scheme like reinforcement learning [4–6].
The free-energy formulation circumvents these problems by
prescribing the policy in terms of free-energy, which encodes
optimal control (Equation 6)"
\item It is difficult to compute with the free energy principle
  in practice, which makes it physiologically less plausible. Auto-encoder theory presented
  \ref{sec:vae} can help to remove these barriers, by making the whole scheme trainable via
  backprop!
  \end{itemize}

>>>>

\subsection{The right and left TPJ: Prediction error signaling and world semantics}
Finally,
the right and left TPJ are known to have hemispheric differences
according to their cytoarchitectonic borders and gyrification pattern
\citep{seghier2013angular}.
Neuroscientific investigations on hemispheric functional specialization
have highlighted the right versus left cerebral hemisphere as dominant for
attention versus language functions.
%
The anterior RTPJ is a key node of the DMN that  is central for
action initiation during externally structured tasks and
sensorimotor control by integrating supramodal stimulus-guided attention
\citep{corbetta2002control}.
Involvement of this DMN node was repeatedly reported in
multi-step action execution \citep{hartmann2005takes},
visuo-proprioceptive conflict (Balslev et al., 2005) and
multi-modal detection of sensory changes across
visual, auditory, or tactile stimulation in a multi-modal fMRI study
\citep{downar2000multimodal}.
In humans, direct electrical stimulation of the
RTPJ during neurosurgery was associated with altered perception
and stimulus awareness \citep{blanke2002neuropsychology}.
%
Importantly, the RTPJ has been shown to be intimately related to
supramodal prediction and error signaling.
It was argued that the RTPJ encodes actions and ensuing outcomes
without necessarily relating those to outcome value
\citep{liljeholm2013neural, hamilton2008action,
jakobs2009effects}.
Neural activity in the RTPJ has been argued to be responsible
for stimulus-driven attentional reallocation to
salient and surprising sources of information
as a ‘‘circuit breaker’’ that recalibrates control and maintenance systems
\citep{bzdok2013tpj, corbettashul2008}.
% (Bzdok et al., 2013; Corbetta et al., 2008).
Indeed, patients with right TPJ damage have particular difficulties
with multistep actions \citep{hartmann2005takes}.
In the face of large discrepancies between actual and previously predicted
environmental events the RTPJ acts a potential switch between
externally-oriented mind sets focussed on the
sensory world and internally-oriented mind sets focussed
on self-relevant mind wandering through mental scenarios.
Transient RTPJ in humans for instance disrupted the
impact of predicted intentions of other individuals
\citep{young2010disruption},
a capacity believed to be subserved by the DMN.
The RTPJ might hence be an important player that shifts away
from the ‘‘internally directed’’ baseline processes
to deal instead with immediate environmental objects and contexts.
% (TODO: slow endogeneous changes in state transitions by the policy and value matrices)

The left TPJ in turn exhibits a close topographical relationship to
Wernicke's area
involved in the comprehension or understanding of written and spoken language.
Neurological patients with damage caused to Wernicke's area
have a major impairment of language comprehension
when listening to others or reading a book.
Their speech
preserves natural rhythm and about normal syntax, yet the
voiced sentences are devoid of meaning (i.e., aphasia).
Abstracting from the typical semantic interpretations in linguistics
and neuropsychology,
the LTPJ probably mediates access to and integrating of world knowlege
to model action concepts
\citep{binder2011neurobiology, seghier2013angular}.
For instance, LTPJ lesions also entail problems in recognizing
others' pantomimed action towards objects
without obvious relation to processing any language content
\citep{varney1987locus}
% (Varney and Damasio 1987; Rothi et al. 1991; Buxbaum et al. 2005).
%
Also inner speech hinges on knowledge retrieval of statistical structure
about the physical and inter-personal world.
Indeed,
the internal production of
formulated thought ("language of the mind") was closely related to the left TPJ
in a multivariate analysis of brain volume
\citep{geva2011neural}.
Further,
episodic memory recall and imagination strongly draw on
complex world knowledge.
Isolated building blocks of world statistics probably get reassembled
in internally generated visual scenarios that
navigate present action, weigh hypothetical possibilities, and forcast the future.
%
The LTPJ may hence facilitate the automated prediction of events
by incorporating experience-derived models of the world
into ongoing action, planning, and problem solving.
% world -> policy matrix driven computation
% self -> value matrix driven computation

\section{Reinforcement Learning, a computational model for DMN function}
We now argue that the DMN is equipped with sufficient neurobiological substrates gradient
for implementing all the major components (memory, reward signaling, accumulators, random sampling, etc.)
of an efficient \textit{reinforcement learning (RL)} algorithm, and all these
components integrate in a plausible way.
Reinforcement learning is learning to achieve a goal by interacting with an external
\textit{environment}. In a general RL framework, an \textit{agent}
interacts with an  environment in a trial-by-error fashion by taking
actions $a$, which trigger changes in the
\textit{state} of the environment
$s \rightarrow s'$, accompanied by a real-valued \textit{reward}
$r = r(s, a)$ (or regret!, since it can be negative) collected by the
agent. In this view, the environmental is partially controlled by
the action of the agent and the reward should be thought
of as an instantaneous satisfaction accompanying the execution of
an action, and may be delayed, only becoming substantial when accumulated over an extended period of time.
%% Here the subscript $t$ denotes time, and can be discrete or
%% continuous.
The environment is generally taken as stochastic,
is \textit{partially observed},
in the sense that only part of the current state is observed by
the agent. This unpredictability
% (in a first-order sense)
is realistic in a model which sets out
to explain DMN function. Indeed, it is known that, the more the external world is predictable,
the more mental activity becomes detached from the actual sensorium
\citep{antrobus1966studies, pope1978regulation}.
Conversely, the more the currently executed task is unknown and unpracticed,
the lower the DMN activity, i.e the tendency for stimulus-independent thoughts
\citep{filler1973daydreaming, teasdale1995stimulus}.
%
These ``offline'' processes may however contribute to optimized control of the organism.
A DMN framework incorporating reinforcement learning
can naturally embed human behavior
into the tension between exploitative action with immediate gains and
explorative action with longer-term reward schedules
\citep{dayan2008decision}.
The implication of the DMN in a diversity of human cognitive processes
can be parsimoneously recast as prediction error minimization
based on pervasive probabilistic simulations,
thus maximizing reward outcome at various time scales.
Such a purposeful optimization objective
may be solved by a stochastic approximation
based on a brain implementation of Markov Chain Monte Carlo sampling
\citep{tenenbaum2011grow}.

\subsection{Partially Observable Markov Decision Processes}
It is our view that at a sufficiently coarse scale, the brain is a physical system governed by the laws of
\textbf{XXX the paper is about the DMN more specifically, perhaps try to sharpen}
physics and can be locally described by Markov processes. Indeed, as noted in \cite{tegmark2016improved},
any system obeying the laws of classical physics can be accurately modeled as a Markov process as long as the time
step is sufficiently short, defining $s(t)$, the state at time $t$, as the position in phase space.
If the process has $memory$ such that the next state depends not only on the current state but also on some finite
number of past states,
rational probabilistic planning can reformulated
as a standard memoryless Markov process by simply expanding the
definition of the state $s$ to include elements of the past.

In artificial intelligence (AI), a popular abstract model for
multi-step decision processes in such an environment are so-called
\textit{Partially Observable Markov Decision Processes (POMDPs)}.
A POMDP models a decision process in which it is assumed that environment dynamics are determined by a Markov process,
but the agent cannot directly observe the underlying state. Instead, it tries to optimize a \textit{subjective} reward
signal which is tried alongside each transition in the environments state-space, and maintains a probability distribution over the set of possible
states, based on a set of observations and observation probabilities. This is a minimal amount of assumptions that can be made about an environment,
and is characteristic of so-called \textit{model-free RL}, the approach adapted by our work.
In contrast, \textit{model-based RL} starts off with some mechanistic assumptions about the dynamics of world.
These can be assumptions about the physical laws governing the agent's world or constraints on the state-space and transitions thereupon.
For example, if a rat in a maze knows
that standing still will produce no change in the environment, and in particular will not eventually lead to finding the food.
\textbf{XXX last sentence not statistically sound}
\textbf{XXX this paper is mostly about the human DMN, therefore it would be better to rephrase the example into something with a human being}
We formalize
a policy matrix to capture the space of possible actions that can be performed
on the environment given the current state. A value matrix
maps expected rewards to environmental objects and events.

@@

\subsection{Value mapping and action policies}
The behavior of the agent is governed by some kind of \textit{policy},
which maps states of the world to
a set of candidate actions to perform given this state. Starting a time $t=0$,
a policy $\pi$ generates a trajectory as follows
\begin{eqnarray*}
  \begin{split}
    S_1 &\sim p(.|S_0,A_0), \text{ collect rewards }R_0 = r(S_0, A_0)\\
    A_1 &\sim \pi(.|S_1)\\
    S_2 &\sim p(.|S_1,A_1), \text{ collect rewards }R_1 = r(S_1, A_1)\\
    \ldots\\
    A_{t} &\sim \pi(.|S_{t})\\
    S_{t+1} &\sim p(.|S_{t},A_{t}), \text{ collect rewards }R_{t} = r(S_{t}, A_{t})\\
    \ldots
  \end{split}
\end{eqnarray*}
Since an action taken now might have repercussions long into the feature, it turns out that the
quantity to optimize is not the instantaneous rewards $r(s, a)$, but a
cumulative reward estimate which takes into account the future. A standard for
modeling this accumulation is the so-called
time-discounted \textit{cumulative reward} function
\begin{equation}
  \label{eq:cumr}
  G^\pi = \sum_{t=0}^{\infty}\gamma^{t}R_t.
\end{equation}
This random variable\footnote{Random as it depends both on the environment's dynamic and the
  policy $\pi$ being played (which can be stochastic).}  measures the cumulative reward of
following a policy $\pi$.
\begin{mdframed}
  \textbf{Where is ``value'' buffered in the brain ? You mean where in the DMN!
  then the answer is in the vmPFC which is connected to the NAc}
  The amygdala is known to be involved in significance evaluation) and
  the NAc is involved in reward evaluation.
\end{mdframed}
  The  goal of RL is then adapt this policy so as to  maximize the cumulative rewards. In
  the definition of $G^\pi$ above, the constant $\gamma$ $(0 \le \gamma < 1$) is the reward \textit{discount factor}.
Taking $\gamma = 0$   corresponds to a perfectly myopic agent who is solely concerned about their
immediate rewards and has no horizon for the future, which is not compatible with long-term goal planning
as potentially implemented in the DMN. For there to be any learning at all, it is thus necessary that $0 < \gamma < 1$.
% \footnote{In which case we can take $T \rightarrow +\infty$, provided the reward function $r$ is bounded.}.
Thus, the agent is positively biased towards considering rewards which with
a short time delay. The reward importance hence decays exponentially
with distance in the future,
given that there is more uncertainty in the farsighted future.
The reward discount factor $\gamma$ can thus be seen as calibrating
the delay discounting behavior of the intelligent agent,
that is the decision related to reward-delay tradeoff schedules.
It is also important to appreciate that a stochastic policy estimation
is more advantageous in many RL settings.
For instance, a deterministic strategy in playing rock-paper-scissors can
be easily exploited and a uniformly random choice is optimal.
\textbf{XXX I love this example :-)}

The agent's goal can then be formulated as finding policy $\pi$ which maximizes the
expected cumulative value of state-stat-pairs $(A_0,S_0)$, also known as the \textit{Q-value}, namely
\begin{equation}
  \label{eq:qval}
  Q^{\pi}(s,a) = \mathbb E [G^\pi|S_0=s,A_0=a,A_1 \sim \pi,A_2 \sim \pi,\ldots].
\end{equation}
\begin{mdframed}
In plain English, the Q-value $Q^{\pi}(s,a)$ is the expected future reward
\textbf{XXX let's stick to "expected" in these situations because
it implies that the DMN is doing prediction without needing to 
emphasize it too much all the time}
over all possible
\textbf{XXX action???}
trajectories in which the agent starts off starts at with an environment in state
$s$, takes and action $a$, and then follows the policy $\pi$ thereafter to select future actions.
\end{mdframed}
%% Given that each $\pi$ correspondence to an exhaustive description of
%% behavioral patterns, $Q^{\pi}(S_t,A_t)$ corresponds to the epistemic value of
%% the set of actions given states in finite time.
By construction, maximizing the expectation in  \eqref{eq:qval} is intractable in general
\textbf{XXX coputationally or statistically?}
due to the enormous number of possible
\textbf{XXX action???}
trajectories the agent might need to
consider.
% \begin{mdframed}
%   \textbf{N.B.:}
%   If the Bellman-based computational ideas have a corresponding
%   neurobiological implementation, this provides a framework of how
%   optimal behavior could be computable in the human brain.
% \end{mdframed}
One popular simplification that removes
%this issue
\textbf{excess in complexity on the agent's horizon???}
is to only consider the sub-family of
\textit{deterministic policies}, which map each state to the best single best action which
should be taken at thin state, and therefore define functions from states to actions.
\textit{Q-learning} (Watkins \& Dayan, 1992), a commonly used off-policy algorithm, uses the
\textit{greedy policy}
\begin{equation}
  \pi(s) = \argmax_{a \in \mathcal A}{Q}(s, a).
  \label{eq:qlearning}
\end{equation}
% where $\tilde{Q}(s, a)$ is the current estimate of the state-action value
% table.
The \textit{Bellman's equation} \cite{sutton1998reinforcement} then takes the simple form
\begin{equation}
  \begin{split}
    Q^\pi(s, a) &= \mathbb E_{s' \sim p(.|s,a)} [r(s,a) + \gamma \max_{a' \in \mathcal A}Q^\pi(s', a')]
    \\
    &= \mathbb E_{s' \sim p(.|s,a)} [r(s,a) + \gamma Q^\pi(s', \pi(s'))],
  \end{split}
  \label{eq:bellman}
\end{equation}
a \textit{fixed-point} equation which provides a recursive decomposition of optimal behavior.
A complicated dynamic programming optimization can be
broken into simpler sub-problems at different time points,
which has often been closely related to the medial prefrontal part of the DMN
\citep{koechlin1999role, braver2002role}.
Using the Bellman equation, each state can be associated with a certain value,
to guide action towards a better state, thus improving on the agent's current policy.
% It is a fixed-point equation for the deterministic policy $\pi$.
Notice that in the expectation in \eqref{eq:bellman}, the sampling is done over only things which
depend on the environment, and so can be learned off-policy by observing state-transitions
triggered by another behavioral policy, which can be stochastic.

@@

\section{Relation to other statistical models (Elvis)}
\subsection{Friston's Free-energy principle (refer to \ref{label:friston})}
\textbf{XXX: add tons of more examples and connnections}
\textbf{XXX: integrate aspects from section "ThE MARkOv DECISION PROBLEM" in Dayan and Daw}
\textbf{XXX: integrate aspects from WakeNsleep by Hinton}
\textbf{XXX: integrate aspects from Friston2014(Dropbox) on active inference}
\textbf{XXX: integrate aspects from Sutton/Barto book chapter}
In summary:
\begin{itemize}
  \item Friston's free-energy principle: brain as inference engine
  biological systems, including brains, must minimize the long-term average of surprise;
  \item A common point between the Free Energy principle and our framework
  is placing the minimization of suprise at the core of brain function.
  In RL, surprise minimization reduces to accurate prediction of
  rewarding outcomes in the future.
  \item Ortega et al. 2013 \citep{ortega2013thermodynamics} ``Thermodynamics as a theory of decision-making with information-processing costs''
\item Friston's critic of RL \citep{fristonAIorRL} (he proposes "active inference''): "This equation (equation 18) comes from the theory of dynamic programming,
pioneered in the 1950s by Richard Bellman and colleagues. To
optimise control a~pð Þ x~ under this formulation, we have to: (i)
assume the hidden states are available to the agent and (ii) solve
Equation 18 for the value-function. Solving for the value-function
is a non-trivial problem and usually involves backwards induction
or some approximation scheme like reinforcement learning.
The free-energy formulation circumvents these problems by
prescribing the policy in terms of free-energy, which encodes
optimal control (Equation 6)"
\item It is difficult to compute with the free energy principle
  in practice, which makes it physiologically less plausible. Auto-encoder theory presented
  \ref{sec:vae} can help to remove these barriers, by making the whole scheme trainable via
  backprop!
  \end{itemize}
